---
title: "HW #9"
subtitle: "10.3, 10.10, 10.13, 10.19, 10.28, 10.32"
date: "`r Sys.Date()`"
author: "Hayden Atchley"
output:
  bookdown::pdf_document2:
    toc: false
    latex_engine: lualatex
    number_sections: false
mainfont: Gentium Book Basic
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.pos = "H"
)

pacman::p_load(Sleuth3, tidyverse, ggthemes, broom, kableExtra, GGally, default)

options(knitr.kable.NA = '')

default(kbl) <- list(booktabs = TRUE, linesep = "")
default(kable_styling) <- list(latex_options = "HOLD_position")
```

# 10.3

In both `(a)` and `(b)`, the _t_-test cannot be used because there are multiple parameters.

# 10.10

```{r}
f <- tribble(
	~Model, ~RSS, ~DF,
	"Full", 5.99713, 32,
	"No Interactions", 8.38155, 34
) %>% 
	mutate(
		"Mean Sq" = RSS / DF,
		"Extra SS" = c(RSS[2] - RSS[1], NA),
		"Extra DF" = c(2, NA)
	) 

f_stat <- unlist(f[1,5] / f[1,6] / f[1,4])
p_val <- pf(f_stat, unlist(f[1,6]), unlist(f[1,3]), lower.tail = FALSE)

f %>% 
	kbl(digits = c(NA,2,0,3,2,0)) %>% 
	kable_styling()
```

The _F_-statistic is therefore $\frac{`r f[1,5] %>% round(2)` / `r f[1,6]`}{`r f[1,4] %>% round(3)`} = `r f_stat %>% round(2)$, and $P(F_{2,32} > `r f_stat %>% round(2)`) = `r p_val %>% round(3)`$. This gives evidence that the slopes are different for the different species.

# 10.13

```{r}
bats <- case1002 %>% 
	as_tibble() %>% 
	mutate(Type = case_when(
		Type == "non-echolocating birds" ~ "bird",
		Type == "non-echolocating bats" ~ "nbat",
		Type == "echolocating bats" ~ "ebat"
		),
		Type = as.factor(Type))

mod_a <- bats %>% 
	mutate(Type = Type %>% relevel("nbat")) %>% 
	lm(log(Energy) ~ log(Mass) + Type, data = .)

mod_c <- bats %>% 
	mutate(Type = Type %>% relevel("ebat")) %>% 
	lm(log(Energy) ~ log(Mass) + Type, data = .)
```

## a

```{r}
mod_a %>% 
	tidy() %>% 
	kbl(digits = c(NA, 3, 3, 2, 3)) %>% 
	kable_styling()
```

## b

The slope of log(Energy) on log(Mass) is 0.815 for all subjects. The intercept is -1.576 for nbats, -1.576 + 0.079 = -1.497 for ebats, and -1.576 + 0.102 = -1.474.

## c

```{r}
mod_c %>% 
	tidy() %>% 
	kbl(digits = c(NA, 3, 3, 2, 3)) %>% 
	kable_styling()
```

## d

This gives the same slope as in `(b)`, and the intercepts are -1.498 for ebat, -1.474 for birds, and -1.577 for nbat.

## e

The test that the intercepts for ebat and birds are the same is given from the table in `(c)`. This intercept has a _p_-value of 0.883, which gives no evidence that the difference in intercepts is not 0.

# 10.19


