---
title: "HW #7"
subtitle: "7.2, 7.9, 7.15, 7.25, 7.29"
date: "`r Sys.Date()`"
author: "Hayden Atchley"
output:
  bookdown::pdf_document2:
    toc: false
    latex_engine: lualatex
    number_sections: false
mainfont: Gentium Book Basic
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

pacman::p_load(Sleuth3, tidyverse, ggthemes, broom, kableExtra)
```


# 7.2

Improved distance measurements would decrease measurement error, resulting in smaller variation and therefore more precise estimates of the regression coefficients.

# 7.9

The intercept of 5kg does _not_ imply that males of height 0 weigh 5kg on average. This does not render the model meaningless, however, as the model is fitted on a narrow band of height and weight data compared to all theoretical real-number values. A height of 0 is well outside this band, and the model is not designed to predict this value.

# 7.15

```{r}
#| pollen-removal,
#| fig.pos="H"

ex0327 %>% 
	filter(BeeType == "Queen") %>% 
	ggplot(aes(x = DurationOfVisit, y = PollenRemoved)) +
	geom_point() +
	geom_smooth(method = 'lm') +
	theme_pander()
```

And a table of the model fit:

```{r}
ex0327 %>% 
	lm(PollenRemoved ~ DurationOfVisit, data = .) %>% 
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(NA,4,4,2,4)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

# 7.25

A plot of the data:

```{r}
ex0725 %>% 
	ggplot(aes(x = Distance, y = Velocity)) +
	geom_point() +
	geom_smooth(method = 'lm') +
	theme_pander() +
	labs(x = "Distance (megaparsec)", y = "Velocity (km/s)")
```

Fitting the data to a linear model with an intercept term gives:

```{r}
#| big-bang-intercept

ex0725 %>% 
	lm(Velocity ~ Distance, data = .) %>% 
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(0,0,1,2,3)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

This does not give evidence that the intercept is different from 0, so we fit without an intercept term as well:

```{r}
#| big-bang-noint

mod <- ex0725 %>% 
	lm(Velocity ~ Distance - 1, data = .)

mod %>% 
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(0,0,2,2,3)) %>%
	kable_styling(latex_options = "HOLD_position")

slope <- mod$coefficients[1] %>% round(2)
age <- (mod$coefficients[1] * 30856775812799586000 / 31557600000000000) %>% 
	round()
```

Because there is no evidence from these data that the intercept is not equal to zero, and because Hubble's theory predicts that the intercept _should_ be zero, we will use the second model to calculate the age of the universe. This gives us:

\begin{align*}
Age &= `r slope` \frac{\text{megaparsec}}{\text{km} / \text{sec}}
\times \frac{3.086 \times 10^{19} \text{km}}{\text{megaparsec}}
\times \frac{\text{billion years}}{3.156 \times 10^{16} \text{sec}} \\
&= `r age` \text{ billion years}
\end{align*}

This is many orders of magnitude greater than the current scientific estimate of around 14 billion years for the age of the universe

# 7.29

A plot of the data:

```{r}
#| fig.pos = "H"

ex0729 %>% 
	ggplot(aes(x = Distance, y = OverEstimate)) +
	geom_point() +
	geom_smooth(method = 'lm') +
	theme_pander() +
	labs(x = "Distance (ft)", y = "Over-estimate (percentage points)")
```

And a summary of the model:

```{r}
ex0729 %>% 
	lm(OverEstimate ~ Distance, data = .) %>% 
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(NA,3,4,2,5)) %>% 
	kable_styling(latex_options = "HOLD_position") 
```

This gives overwhelming evidence that there is a correlation between over-estimation of Kerry votes and distance of the pollsters from the door, lending credence to the theory that Bush voters were more likely to avoid exit polls.
