---
title: "HW #7"
subtitle: "7.2, 7.9, 7.15, 7.25, 7.29"
date: "`r Sys.Date()`"
author: "Hayden Atchley"
output:
  bookdown::pdf_document2:
    toc: false
    latex_engine: lualatex
    number_sections: false
mainfont: Gentium Book Basic
---

```{r setup}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

pacman::p_load(Sleuth3, tidyverse, ggthemes, broom)
```


# 7.2

Improved distance measurements would decrease measurement error, resulting in smaller variation and therefore more precise estimates of the regression coefficients.

# 7.9

The intercept of 5kg does _not_ imply that males of height 0 weigh 5kg on average. This does not render the model meaningless, however, as the model is fitted on a narrow band of height and weight data compared to all theoretical real-number values. A height of 0 is well outside this band, and the model is not designed to predict this value.

# 7.15

```{r}
#| pollen-removal,
#| fig.pos="H"

ex0327 %>% 
	filter(BeeType == "Queen") %>% 
	ggplot(aes(x = DurationOfVisit, y = PollenRemoved)) +
	geom_point() +
	geom_smooth(method = 'lm') +
	theme_pander()
```

<!-- This fit is not a very good one at all. Very little of the actual data is within the 95% confidence band predicted, and the band is very wide for much of the data range. Some sort of transformation or other model is likely to produce a better result. -->

# 7.25

Fitting the data to a linear model with an intercept term gives:

```{r}
#| big-bang-intercept

ex0725 %>% 
	{lm(Velocity ~ Distance, .)} %>% 
	tidy() %>% 
	knitr::kable()
```

This does not give evidence that the intercept is different from 0, so we fit without an intercept term as well:

```{r}
#| big-bang-noint

ex0725 %>% 
	{lm(Velocity ~ Distance - 1, .)} %>% 
	tidy() %>% 
	knitr::kable()
```


