---
title: "HW #6"
subtitle: "6.4, 6.11, 6.12, 6.16, 6.21, 6.23"
author: Hayden Atchley
date: "`r Sys.Date()`"
mainfont: Gentium Book Basic
output:
  bookdown::pdf_document2:
    number_sections: false
    toc: false
    latex_engine: lualatex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
options(
	knitr.kable.NA = ''
)

pacman::p_load(Sleuth3, tidyverse, kableExtra, broom, ggthemes, emmeans, DescTools)
```

# 6.4

(d) is true: the range cannot be smaller than 2.2 Btu/lb.

# 6.11

The _t_-ratio will be the same since both the estimate and standard error will be scaled by a factor of 3.

# 6.12

We start with a look at the means by group:

```{r}
lsm <- lsmeans(
	lm(Score ~ Handicap, case0601),
	"Handicap")

lsm %>% 
	kbl(booktabs = TRUE, digits = c(0,2,3,0,2,2)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

We want to contrast the _mobility_ handicaps (amputee, crutches, wheelchair) with the _communication_ handicaps (hearing), so we use the linear combination

\begin{align*}
	\gamma &= \frac{A + C + W}{3} - H \\
	&= \frac{1}{3}A + \frac{1}{3}C + \frac{1}{3}W - H.
\end{align*}

This gives us:

```{r}
contrast(lsm, list("Mobility vs. Communication" = c(1/3, 1/3, -1, 0, 1/3))) %>% 
	kbl(booktabs = TRUE, digits = c(0,2,3,0,3,3)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

This indicates that there is good evidence of a difference, though not overwhelming.

# 6.16

We have data with `r (n <- 36)` subjects divided into `r (groups <- 6)` groups. For the _LSD_ method, we use $n - I = 30$ degrees of freedom, so the 97.5% critical value of $t_{30}$ (and the multiplier) is `r qt(0.975, 30) %>% round(3)`. Since the _F_-test _p_-value is large (0.085), using the _F_-protected _LSD_ comparison we would declare no difference between any groups. Tukey-Kramer gives us a multiplier of $\frac{q_{6,30}(0.95)}{\sqrt{2}} = \frac{`r (q <- qtukey(0.95, 6, 30)) %>% round(3)`}{\sqrt{2}} = `r (q/sqrt(2)) %>% round(3)`$. Using Bonferroni we get $k = \frac{6(6-1)}{2} = `r (k <- 6*(6-1)/2)`$, and so we use a multiplier of $t_{30}(1-\frac{0.05}{2k}) = t_{30}(`r (p <- (1 - (0.05/2/k))) %>% round(4)`) = `r qt(p, 30) %>% round(3)`$. Scheffe's multiplier is $\sqrt{(6-1)F_{5,30}(0.95)} = `r sqrt(5*qf(0.95, 5, 30)) %>% round(3)`$.

# 6.21

We start with an ANOVA analysis:

```{r}
data <- ex0525 %>% 
	mutate(Educ = str_replace(Educ, "-", ":"))

model <- aov(Income2005 ~ Educ, data)

model %>% 
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(0,0,0,0,2,2)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

This shows that there are significant differences between groups.

## a

Using Tukey-Kramer:

```{r}
TukeyHSD(model) %>% 
	tidy() %>% 
	select(-term, -null.value) %>% 
	kbl(booktabs = TRUE, digits = c(0,0,0,0,4), linesep = "") %>% 
	kable_styling(latex_options = "HOLD_position")
```

All of these differences are significant except for "12 - <12" and "16 - >16", though "13:15 - 12" is less significant than the rest.

## b

Using Dunnett:

```{r}
DunnettTest(data$Income2005, data$Educ, control = "12") %>% 
	{.$`12`} %>% 
	kbl(booktabs = TRUE, digits = c(0,0,0,4)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

All of these differences are significant except for "<12 - 12".

# 6.23

Looking at a plot of the data:

```{r}
#| fig.pos="H",
#| fig.height=3

ex0623 %>% 
ggplot(aes(WtLoss24, Group)) +
	geom_violin(aes(fill = Group)) +
	stat_boxplot(width = 0.3, fill = NA) +
	geom_vline(xintercept = 0) +
	theme_pander() +
	guides(fill = "none") +
	labs(x = "Weight Loss")
```

Running a one-way ANOVA test gives:

```{r}
model <- aov(WtLoss24 ~ Group, ex0623)

model %>% 
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(0,0,2,2,3,4), linesep = "") %>% 
	kable_styling(latex_options = "HOLD_position")
```

This gives weak evidence that there is a difference between at least two of the means. A Tukey analysis at 95% confidence gives:

```{r}
model %>% 
	TukeyHSD() %>% 
	tidy() %>% 
	select(-term, -null.value) %>% 
	kbl(booktabs = TRUE, digits = c(0,3,3,3,4), linesep = "") %>% 
	kable_styling(latex_options = "HOLD_position")
```

This gives evidence that there is a difference between the Low-Fat and Low-Carb diets, though again, the evidence is not overwhelming.
