---
title: "HW #1"
subtitle: |
  1.2, 1.6, 1.11, 1.12, 1.16, 1.18, 1.26, 1.27
author: Hayden Atchley
date: '`r format(Sys.Date(), "%e %B %Y")`'
output: 
  pdf_document: 
    fig_caption: yes
    latex_engine: lualatex
mainfont: Gentium Book Basic
# mathfont: Gentium Book Basic
fonsize: 11pt
tabstop: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(EnvStats)
library(Sleuth3)
library(knitr)
library(kableExtra)
library(DescTools)
library(Sleuth3)
```

# 2.3
False

# 2.6


# 2.7
The $p$-value for the hypothesis that the mean difference is zero must be less than 0.05.

# 2.13
First I separated the dataset into fish oil and regular oil treatments:

```{r}
fishoil <- ex0112 %>% 
	filter(Diet == "FishOil")
regoil <- ex0112 %>% 
	filter(Diet == "RegularOil")
```

Computing the sample averages and standard deviations:

```{r}
oilcomp <- tibble(
	Diet = c("Fish Oil", "Regular Oil"),
	n = c(length(fishoil$BP), length(regoil$BP)),
	`Sample average` = c(mean(fishoil$BP), mean(regoil$BP)),
	`Sample stdev` = c(sd(fishoil$BP), sd(regoil$BP))
)

oilcomp %>% 
	kbl(booktabs = TRUE, digits = 2, position = "h") %>% 
	kable_styling()
```

```{r, echo=FALSE}
nf <- oilcomp$n[1] %>% round()
nr <- oilcomp$n[2] %>% round()

mf <- oilcomp$`Sample average`[1] %>% 
	round(2)
mr <- oilcomp$`Sample average`[2] %>% 
	round(2)

sdf <- oilcomp$`Sample stdev`[1] %>% 
	round(2)
sdr <- oilcomp$`Sample stdev`[2] %>% 
	round(2)

diff = mf-mr

psd <- sqrt( ((nf-1)*sdf+(nr-1)*sdr) / (nf+nr-2))
pse <- (psd*sqrt(1/nf + 1/nr))
df <- (nf + nr - 2) %>% round()

conf <- 5
ttest <- t.test(fishoil$BP, regoil$BP,
								var.equal = TRUE, alternative = "greater")
tstat <- ttest$statistic
pval <- ttest$p.value
```

The pooled standard deviation is given by: 
\[
\sqrt{\frac{(`r nf`-1)*`r sdf` + (`r nr`-1)*`r sdr`}{`r nf`+`r nr`-2}},
\]
which is equal to `r psd %>% round(2)`, and the pooled standard error of the difference in means is $`r psd` * \sqrt{\frac{1}{`r nf`} + \frac{1}{`r nr`}}$,  which equals `r pse %>% round(2)`. There are $`r nf`+`r nr`- 2 = `r df`$ degrees of freedom associated with this estimate, and the 97.5th percentile of a _t_-distribution with `r df` degrees of freedom is `r qt(0.975, df) %>% round(2)`.

A 95% confidence interval is given by $\mu \pm something$, which in this case gives $`r diff` \pm conf$. The _t_-score is `r tstat %>% round(2)` and the one-sided _p_-value is `r pval %>% round(4)`.

# 2.18
First I separated the data into 1976 and 1978 finches:

```{r}
f6 <- ex0218 %>% 
	filter(Year == 1976)
f8 <- ex0218 %>% 
	filter(Year == 1978)
```


# 2.22