---
title: "HW #5"
subtitle: "5.4, 5.10, 5.16, 5.17, 5.18, 5.23"
author: Hayden Atchley
date: "`r Sys.Date()`"
mainfont: Gentium Book Basic
output:
  bookdown::pdf_document2:
    number_sections: false
    toc: false
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
options(
	knitr.kable.NA = ''
)

pacman::p_load(Sleuth3, tidyverse, kableExtra, broom, ggthemes)
```

# 5.4

A treatment usually is associated with a change in the mean but not the variance, so the best estimate of $\sigma$ is $s_p$ from all the groups.

# 5.10

Fewer degrees of freedom, especially in the denominator, result in distributions where larger _F_-statistics are more likely.

# 5.16

```{r}
aov(Percent ~ Judge, case0502) %>%
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(0,0,1,2,3,8)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

# 5.17

```{r}
windf <- 24
winssq <- 35088
totdf <- 31
totssq <- 70907

betdf <- totdf - windf
betssq <- totssq - winssq

numgroups <- betdf + 1
```

```{r}
tibble(
	Source = c("Between Groups", "Within Groups"),
	DF = c(betdf, windf),
	SumSq = c(betssq, winssq),
	MeanSq = SumSq/DF,
	Fstat = c(MeanSq[1]/MeanSq[2], NA),
	Pval = c(pf(Fstat[1], DF[1], DF[2], lower.tail = FALSE), NA)
) %>% 
	kbl(booktabs = TRUE, digits = c(0,0,0,0,2,4)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

There were `r numgroups` groups, and there is evidence at the $\alpha = 0.01$ level that the group means are _not_ identical.

# 5.18

```{r}
means.treat <- ex0518 %>% 
	group_by(Treatment) %>% 
	summarise(mean.treat = mean(Protein))

means.treatday <- ex0518 %>% 
	group_by(Treatment, Day) %>% 
	summarise(mean.treatday = mean(Protein)) %>% 
	mutate(Treatment = relevel(Treatment, "CPFA50"))

resid_table <- ex0518 %>%
	left_join(means.treat) %>% 
	left_join(means.treatday) %>% 
	mutate(
		resid.treat = Protein - mean.treat,
		resid.treatday = Protein - mean.treatday
	)

model_treat <- aov(Protein ~ Treatment, ex0518)

model_treatday <- aov(Protein ~ Treatment + Day, ex0518)
```

## a

```{r}
means.treat %>% 
	kbl(booktabs = TRUE, digits = 1) %>% 
	kable_styling(latex_options = "HOLD_position")
```

```{r}
#| fig.height=3


model_treat %>% 
	augment() %>% 
	ggplot(aes(x = .fitted, y = .resid)) +
	geom_point() +
	theme_pander() +
	labs(x = "Fitted", y = "Residuals")
```

```{r}
#| fig.height=3

resid_table %>% 
	ggplot(aes(x = Day, y = resid.treat)) +
	geom_point() +
	theme_pander() +
	labs(x = "Day", y = "Residuals")
```

## b

```{r}
means.treatday %>% 
	arrange(Treatment, Day) %>% 
	pivot_wider(names_from = Day, values_from = mean.treatday) %>% 
	kbl(booktabs = TRUE, digits = 1) %>% 
	kable_styling(latex_options = "HOLD_position")
```

```{r}
aov(Protein ~ TrtDayGroup, ex0518) %>% 
	tidy() %>%
	kbl(booktabs = TRUE, digits = c(0,0,1,2,3,5)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

From this we can conclude that the means of all 10 groups are _not_ equal.

## c

We compare the two models as follows:

```{r}
anova(model_treat, model_treatday) %>%
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(0,0,1,0,1,3,3)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

This gives strong evidence that the means between days are different.





# 5.23

A boxplot of the data is given below:

```{r}
#| fig.pos="h",
#| fig.height=2.8

ex0523 %>% 
	mutate(Bone = str_remove(Bone, "Bone"),
				 Bone = reorder(Bone, sort(as.integer(Bone)))) %>% 
	ggplot(aes(x = Bone, y = Oxygen, fill = Bone)) +
	geom_boxplot() +
	theme_pander() +
	guides(fill = "none")
```

The results of an ANOVA test are given below:

```{r}
aov(Oxygen ~ Bone, ex0523) %>% 
	tidy() %>% 
	kbl(booktabs = TRUE, digits = c(0,0,2,3,2,7)) %>% 
	kable_styling(latex_options = "HOLD_position")
```

Based on these results, there is strong evidence against the means being equal.

