---
title: "Homework #3"
subtitle: |
  3.3, 3.5, 3.8, 3.18, 3.26, 3.28, 3.30
author: "Hayden Atchley"
date: '`r format(Sys.Date(), "%e %B %Y")`'
mainfont: Gentium Book Basic
output:
  pdf_document:
    toc: false
    fig_caption: yes
    latex_engine: lualatex
fonsize: 11pt
tabstop: 2
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(Sleuth3)
library(ggeasy)
library(tidyverse)
library(ggthemes)
library(scales)
```

# 3.3
If the crew were aware of whether seeding was conducted, they may have---consciously or not---changed their behaviors in a way that would influence the results.

# 3.5
Though there are several samples of dioxin levels in the Vietnam veterans that are substantially higher that in other veterans, these are all extreme outliers. Additionally, there are over 6 times as many observations in the former group, so the outliers represent an even smaller proportion of this group. Further analysis is required to determine if the difference is stastically significant.

# 3.8
The observations are not independent, but that is by design. The paired _t_-test is used in this case, since each observation is paired with another, related observation.

# 3.18

## a
The difference in means is _not_ resistant, as a dramatic change in one observation changes this statistic considerably.

## b
The average of the ranks for the trauma group minus the average of the ranks for the non-trauma group _is_ resistant, as a dramatic change in one observation would not change the rankings, and therefore the difference of their averages, very much if at all.

# 3.26
The log-transformed data from the Agent Orange study (adding 0.1 to all values since the dataset contains zeros) gives:

```{r, fig.pos="h"}
logdata <- case0302 %>% 
	mutate(add = Dioxin + 0.1,
				 log = log(add))

logdata %>% 
	ggplot(aes(y = Veteran, x = log)) +
	geom_boxplot() +
	theme_gdocs() +
	labs(x = "log(Dioxin Level)")
```

```{r}
splitdata <- logdata %>% 
	split(.$Veteran)

test <- t.test(splitdata$Vietnam$log, splitdata$Other$log)

p <- test$p.value %>% round(3)
low <- test$conf.int[1] %>% round(3)
high <- test$conf.int[2] %>% round(3)

lowexp <- exp(low) %>% round(3)
highexp <- exp(high) %>% round(3)
```

Running a _t_-test on these transformed samples gives a _p_-value of $`r p`$ and a 95% confidence interval of $`r low` \less \mu_2 - \mu_1 \less `r high`$ for the log-transformed data. Back-transforming gives us that the median dioxin of the Vietnam veterans is between $e^{`r low`} = `r lowexp`$ and $e^{`r high`} = `r highexp`$ times as much as that of the other veterans.

# 3.28
A boxplot of the data:

```{r, fig.pos="h"}
sparrows <- ex0221 %>% 
	split(ex0221$Status)

ex0221 %>% 
	ggplot(aes(y = Status, x = Humerus)) +
	geom_boxplot() +
	theme_gdocs() +
	labs(x = "Humerus Length", y = "")
```

```{r}
perish_min <- min(sparrows$Perished$Humerus)

no_small <- ex0221 %>% 
	filter(!(Status == "Perished" & Humerus == perish_min))
no_small <- no_small %>% 
	split(no_small$Status)

test_all <- t.test(sparrows$Survived$Humerus, sparrows$Perished$Humerus)
test_nosmall <- t.test(no_small$Survived$Humerus, no_small$Perished$Humerus)

p_all <- test_all$p.value %>% round(3)
p_nosmall <- test_nosmall$p.value %>% round(3)
```

A _p_-value for the difference in humerus lengths with all observations is `r p_all`, and without the smallest observation in the perished group (`r perish_min`), the _p_-value is `r p_nosmall`. The _p_-value changes a fair bit (by `r p_nosmall - p_all`), but even with all the observations included, there is not strong evidence that the difference in means is not equal to zero. If there were a more drastic difference such that one of these tests _did_ show strong evidence of a difference in means, then the removed observation should be double-checked for accuracy.

# 3.30
```{r}
income <- ex0330 %>% 
	select(-Subject) %>% 
	mutate(Educ = as.factor(Educ))
```

A boxplot of the data:

```{r}
income %>% 
	ggplot(aes(y = Educ, x = Income2005)) +
	geom_boxplot() +
	theme_gdocs() +
	labs(x = "Income", y = "Years of Education") +
	scale_x_continuous(labels = label_comma())
```

This data is highly skewed in the positive direction, so we look at a log transform as well:

```{r}
incomelog <- income %>% 
	mutate(log = log(Income2005))

incomelog %>% 
	ggplot(aes(y = Educ, x = log)) +
	geom_boxplot() +
	theme_gdocs() +
	labs(x = "log(Income)", y = "Years of Education") +
	scale_x_continuous(labels = label_comma())

split_income <- incomelog %>% 
	split(incomelog$Educ)
```

```{r}
test <- t.test(split_income$`16`$log, split_income$`12`$log)

p <- test$p.value %>% round(3)
low <- test$conf.int[1] %>% round(3)
high <- test$conf.int[2] %>% round(3)
diff <- (test$estimate[1] - test$estimate[2]) %>% round(3)
delta <- ((high - low)/2) %>% round(3)
```

Running a _t_-test on this transformed data gives a _p_-value of `r p`, with a 95% confidence interval of `r diff` $\pm$ `r delta`. Restating in terms of the original data, the median income for those with 16 years of education is between $e^{`r low`} = `r exp(low) %>% round(2)`$ and $e^{`r high`} = `r exp(high) %>% round(2)`$ times as much as those with 12 years of education.

